Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-12-14 article.qmd
Add for-loop example in the introduction as a complex extensibility example

To overcome these limitations,
we developed poincaré and SimBio,
open-source Python packages for defining and simulating systems.
Poincaré allows defining differential equation systems,
while SimBio builds on it for defining reaction networks.
They are focused on providing an ergonomic experience to end-users by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
Since they are coded in Python,
every part from model definition to simulation can be extended from Python syntax.
Being the first-ever pure Python packages for systems modelling,
they offer extensive extensibility,
from simple tasks like reusing integrators defined in other packages,
to complex ones like altering the compilation process to leverage some structure in the equations.
**TODO: For example...**
The models built using these packages can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Furthermore,
they have a modular architecture with a clear separation of concerns,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
We showcased the reliability of these tools by benchmarking them against the simulation results from other tools.
We also highlighted the substantial performance improvements our tools offer,
as this is crucial for construction and simulation of models of whole cells and organisms,
which necessitate the simulation of significantly large-scale models.

>>>>

To overcome these limitations,
we developed poincaré and SimBio,
open-source Python packages for defining and simulating systems.
Poincaré allows defining differential equation systems,
while SimBio builds on it for defining reaction networks.
They are focused on providing an ergonomic experience to end-users by integrating well with IDEs and static analysis tools
through the use of standard modern Python syntax.
Since they are coded in Python,
every part from model definition to simulation can be extended from Python syntax.
Being the first-ever pure Python packages for systems modelling,
they offer extensive extensibility,
from simple tasks like reusing integrators defined in other packages,
to complex ones like altering the compilation process to leverage some structure in the equations.
For example,
using a for-loop in the compiled equations could improve the runtime performance
if there is some repetitive structure in the system,
as happens in spatial modelling.
The models built using these packages can be introspected to create other representations,
such as graphs connecting species and/or reactions,
or tables with parameters or equations.
Furthermore,
they have a modular architecture with a clear separation of concerns,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
We showcased the reliability of these tools by benchmarking them against the simulation results from other tools.
We also highlighted the substantial performance improvements our tools offer,
as this is crucial for construction and simulation of models of whole cells and organisms,
which necessitate the simulation of significantly large-scale models.
